A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-01-19 ICLP Paper/samplepaper.tex
progress on intro

>>>>

While decision versions of scheduling problems can be successfully modeled and
solved by an extension of ASP with Difference Logic (DL) constraints
\citep{gebser2016theory}, implemented by \clingodl on top of the
(multi-shot) ASP system \clingo \citep{gekakasc17a},
the optimization capacities of \clingodl come to their limits on moderately sized
yet highly combinatorial JSP instances \citep{elkgeb20a}, for some of which optimal solutions are so far unknown \citep{shysha18a}.
In this paper, we greatly extend our preliminary study \citep{elkgeb20a} 
on problem decomposition into time windows and successive schedule optimization
by means of multi-shot ASP modulo DL solving with \clingodl.
The goal of the decomposition is to split highly complex scheduling tasks into
balanced portions for which partial schedules of good quality can be reliably
found within tight runtime limits, and then be merged into a global solution of
significantly better quality than obtainable in similar runtime 
with single-shot optimization on the full problem.
In this process, problem decomposition must respect the precedence of operations
within their encompassing jobs to guarantee the feasibility of schedules, and as
a secondary objective, the order by time windows of operations sharing a machine should come close to the sequences of operations in (unknown) optimal schedules.
We address computational efficiency as well as solution quality by devising and investigating decomposition strategies regarding the size of
time windows and heuristics to choose their operations.

The contributions of our work going beyond the study in \citep{elkgeb20a} are:
\begin{enumerate}
\item In addition to problem decomposition based on the earliest starting times
      of operations, we consider the most total work remaining criterion and
      refinements of both strategies by bottleneck machines.
      We encode static as well as dynamic decomposition variants, the latter taking
      partial schedules for operations of previous time windows into account, by stratified
      ASP programs (without DL constraints).
\item Considering that a decomposition into time windows may be incompatible with
      the optimal sequences of operations sharing a machine, we incorporate
      time window overlapping into the iterative scheduling process to
      offer a chance for revising "decomposition mistakes".
      Moreover, as the makespan objective we apply for the optimization of
      (partial) schedules tolerates unnecessary idle times of machines as long as
      they do not stretch the scheduling horizon,
      we devise a stratified ASP encoding to postprocess and compress partial
      schedules by reassigning operations to earlier idle times of their machines.
\end{enumerate}

% From a computational perspective, decomposition aims to split highly complex scheduling tasks into better manageable subproblems with a balanced number of 
% operations, so that good-quality or even optimal partial solutions can be
% reliably found in a small fraction of runtime.
% Regarding the feasibility and quality of solutions, problem decomposition must
% respect the precedence of operations within their jobs, and partial schedules
% optimized by time windows should yield better global solutions than obtainable
% in similar runtime on the full problem.
% We devise and investigate a variety of decomposition strategies in terms of the number and size of time windows as well as heuristics for choosing their operations.
% Moreover, we incorporate time window overlapping and compression techniques
% into the iterative scheduling process in order to counteract limitations of window-wise optimization restricted to partial schedules.
% Our experiments on JSP benchmark sets of several sizes show that
% successive optimization by multi-shot ASP solving leads to substantially better schedules within the runtime limit than global optimization on the full problem,
% where the gap increases with the number of operations to schedule.
